---
title: "Oak Data Exploration"
author: "Jessica M Zamudio"
format: html
editor: visual
---

### Install Packages, Library functions

```{r}
#install.packages("tidyverse")
#install.packages("plyr")
library(tidyverse)
library(readxl)
library(plyr)
library(ggplot2)


Survival_HeightApril26Blocks7_6Clean <- read_excel("Oak Project/Data/Survival_HeightApril26Blocks7&6Clean.xlsx", 
                                                   sheet = "Input1", skip = 1,
                                                   na="NA")
View(Survival_HeightApril26Blocks7_6Clean)

names(Survival_HeightApril26Blocks7_6Clean)

```

### Renaming columns of April data sheet

```{r}
#renaming the columns, add dyplr to rename because for some reason it won't work without it
Survival_HeightApril26Blocks7_6Clean <-
  Survival_HeightApril26Blocks7_6Clean |> 
  dplyr::rename(Page_Num ="Page #",
         Date ="MM/DD/YYYY", 
        Recorder ="Name",
       Block  ="#...4",
      Tag_Num   ="#...5",
       Oak_type  ="NR, S, or A",
        Treatment ="AC, OA, NP, DT, or DL",
     OA_microclimate    ="TH, TI, CN",
       Seedling_Ht_cm  ="(cm)",
        Num_leaves ="#...10",
       Num_dead_leaves  ="#...11",
        Leaf_greenness ="%",
        Num_germinated ="#...13",
       RCD  = "(mm)",
       Notes  ="as written on data sheet"
         )
```

### Convert columns to numeric

```{r}
#mutating the character columns to numeric
glimpse(Survival_HeightApril26Blocks7_6Clean)
Survival_HeightApril26Blocks7_6Clean <-
Survival_HeightApril26Blocks7_6Clean |> 
  mutate(Seedling_Ht_cm=as.numeric(Seedling_Ht_cm)) |> 
  mutate(Num_leaves=as.numeric(Num_leaves)) |> 
  mutate(Num_dead_leaves=as.numeric(Num_dead_leaves)) |> 
  mutate(Leaf_greenness=as.numeric(Leaf_greenness)) |> 
  mutate(Num_germinated=as.numeric(Num_germinated)) |> 
  mutate(RCD=as.numeric(RCD))



```

```{r}


```

### Import remaining Oak files

```{r}
#importing the oak data files into session

oak_data <- list.files(path="Oak Project/Data",
                       
                       pattern="*6Clean.xlsx",
                       
                       full.names=TRUE) |> 
  
  lapply(read_excel,
         
         na="NA",sheet="Input1", skip=1 ) |> 
  bind_rows()

oak_data

view(oak_data)
```

### Rename Columns of Oak files

```{r}
oak_data <-
  oak_data |> 
  dplyr::rename(Page_Num ="Page #",
                Date ="MM/DD/YYYY", 
                Recorder ="Name",
                Block  ="#...4",
                Tag_Num   ="#...5",
                Oak_type  ="NR, S, or A",
                Treatment ="AC, OA, NP, DT, or DL",
                OA_microclimate    ="TH, TI, CN",
                Seedling_Ht_cm  ="(cm)",
                Num_leaves ="#...10",
                Num_dead_leaves  ="#...11",
                Leaf_greenness ="%",
                Num_germinated ="#...13",
                RCD  = "(mm)",
                Notes  ="as written on data sheet"
  )
```

```{r}
glimpse(oak_data)
oak_data <-
  oak_data |> 
  mutate(Seedling_Ht_cm=as.numeric(Seedling_Ht_cm)) |> 
  mutate(Num_leaves=as.numeric(Num_leaves)) |> 
  mutate(Num_dead_leaves=as.numeric(Num_dead_leaves)) |> 
  mutate(Leaf_greenness=as.numeric(Leaf_greenness)) |> 
  mutate(Num_germinated=as.numeric(Num_germinated)) |> 
  mutate(RCD=as.numeric(RCD))


```

```{r}
#splitting tag number and acorn number into separate columns
oak_data_CLEAN<- 
  oak_data |> 
  separate(Tag_Num, into = c("Tag_Num", "Acorn_Num"), sep="A") |> 
  mutate(Tag_Num=as.numeric(Tag_Num),
         Acorn_Num=as.numeric(Acorn_Num)) |> 
  mutate(Leaf_greenness=as.numeric(Leaf_greenness),
         Dead = case_when(
           Seedling_Ht_cm==0~"Dead",
           Seedling_Ht_cm>0~"Alive"
         ))

View(oak_data_CLEAN)


```

```{r}
#seedling height
#seedling_summary <- 
  #oak_data_CLEAN |>
 # select(Treatment, OA_microclimate, Seedling_Ht_cm) |> 
 # group_by(OA_microclimate) |> 
 # summarize(
        #    n= n(),
         #   MeanSeedHeight = mean(Seedling_Ht_cm))

#ggplot(OA_microclimate)
#aes(x = Microclimate)
#geom_bar()
```

```{r}
#oak_data_CLEANER<-
#oak_data_CLEAN |> 
 # select(Tag_Num, Acorn_Num) |> 
  #distinct() |> 
  #arrange(Tag_Num)

#view(oak_data_CLEANER)

#Acorn_ID_Table<- oak_data_CLEAN |> 
#    select(Tag_Num) |> 
#    arrange(Tag_Num) |> 
#  mutate(Acorn1=1, Acorn2=2, Acorn3=3) |> 
#  pivot_longer(
#    cols=c(Acorn1, Acorn2, Acorn3),
#    names_to = "Whatever",
#    values_to = "Acorn_Num" |> 
#      select(-Whatever))
```

Q. Effect of treatment:

All 4 responses:

-Seedling height

- focus on focal

- acorn vs NR analyze separate

  - acorn: at end

  - NR: over time

- exclude NA

-Seedling greenness

- focus on focal

- acorn vs NR analyze separate

  - acorn: at end

  - NR: over time

- exclude NA

-germination rate: each tag once, last time only, exclude NRs, only Acorns

-seedling survival: over time OR at end, exclude NAs (non germinates), ?clean data more, summarize to RATE OR logistic regression

!!!!Next assignment: Germination rate for microclimate

### Germination Rate

```{r}
#treatement is that DT AC OA thing
#cleaning germination rate data: each tag once?, last time only?, exclude NRs only Acorns NOT OCT 2025
unloadNamespace("plyr")

#created a data set for germination rate and cleaned data: selected only necessary columns, dropped all values of oak type labeled as NR because we are only looking at acorns and seedlings, looking at only plants that actually grew

germ_rate_data <-
  oak_data_CLEAN |> 
    select(Date, Tag_Num, Acorn_Num, Oak_type, Treatment, OA_microclimate, Seedling_Ht_cm, Num_germinated) |> 
  filter( Oak_type != "NR") |> 
  filter(Seedling_Ht_cm > 0) |> 
  filter( Oak_type != "S")

#organizing dates to get only the latest data entry for each tag number and acorn entry

germ_rate_data_latest <-
  germ_rate_data|>
  group_by(Tag_Num, Acorn_Num) |> 
  arrange(desc(Date)) |> 
  slice(1) |> 
  ungroup()

#calculating the germination rate with the latest entries
germination_rate <-
  germ_rate_data_latest |> 
  group_by(Tag_Num, Acorn_Num) |>
  mutate(germ_rate = (Num_germinated/3)*100)
  
```

### Germination Rate for Acorns

```{r}
#germination rate NA to O value
germination_rate |> 
  mutate(
    germination_rate$germ_rate[is.na(germination_rate$germ_rate)]<- 0
  
    
#Organizing dead and germinated rate into bar graph for treatment types:

#replacing NA values with "dead" 
dead_germ_data<-
  germination_rate|> 
  mutate(
    germ_rate = replace_na(as.character(germ_rate),"Dead"))

#plotting the dead trees with the treatment as fill in bar graph
dead_germ_treatment<-
dead_germ_data |> 
  ggplot(aes(germ_rate, fill= Treatment))+
  geom_bar()
plot(dead_germ_treatment)
  
#organizing dead and germinated rate of microclimate of OA treatment

# first select only OA treatment, then plot germ rate with microclimate 

OA_microclimate_germ<-
  dead_germ_data |> 
  drop_na(OA_microclimate) |> 
  ggplot(aes(germ_rate, fill= OA_microclimate))+
  geom_bar()
plot(OA_microclimate_germ)

#after seeing the results, did they put NA for microclimate if the treat was dead, if so why? we could still analyze what that data means

#plotting open area trees that are all over 0cm, this one does not work for some dumb reason
OA_Death_Microclimate<-
dead_germ_data |> 
  filter(Treatment == "OA") |> 
  ggplot(aes(fill= OA_microclimate))+
  geom_bar()


```

### Germination Rate and Micro-climate data exploration

```{r}
#looking at germination rate trends in a bar graph? helpful? idek. This is acorns only, when fill=Oak_type only Acorn present
germination_rate |> 
ggplot(aes(germ_rate, fill = Treatment)) +
  geom_bar()

#looking at germination rate trends in a bar graph? helpful? idek. acorns only, where fill is the germinations rate and they are separated by treatment type
gr_height<- 
  ggplot(germination_rate, aes(x= germ_rate, y= Seedling_Ht_cm))+
           geom_point()
plot(gr_height)
```
