---
title: "Oak Data Exploration"
author: "Jessica M Zamudio"
format: html
editor: visual
---

### Install Packages, Library functions

```{r}
#install.packages("tidyverse")
#install.packages("plyr")
library(tidyverse)
library(readxl)
library(plyr)
library(ggplot2)


Survival_HeightApril26Blocks7_6Clean <- read_excel("Oak Project/Data/Survival_HeightApril26Blocks7&6Clean.xlsx", 
                                                   sheet = "Input1", skip = 1,
                                                   na="NA")
#View(Survival_HeightApril26Blocks7_6Clean)

names(Survival_HeightApril26Blocks7_6Clean)

```

### Renaming columns of April data sheet

```{r}
#renaming the columns, add dyplr to rename because for some reason it won't work without it
Survival_HeightApril26Blocks7_6Clean <-
  Survival_HeightApril26Blocks7_6Clean |> 
  dplyr::rename(Page_Num ="Page #",
         Date ="MM/DD/YYYY", 
        Recorder ="Name",
       Block  ="#...4",
      Tag_Num   ="#...5",
       Oak_type  ="NR, S, or A",
        Treatment ="AC, OA, NP, DT, or DL",
     OA_microclimate    ="TH, TI, CN",
       Seedling_Ht_cm  ="(cm)",
        Num_leaves ="#...10",
       Num_dead_leaves  ="#...11",
        Leaf_greenness ="%",
        Num_germinated ="#...13",
       RCD  = "(mm)",
       Notes  ="as written on data sheet"
         )
```

### Convert columns to numeric

```{r}
#mutating the character columns to numeric
glimpse(Survival_HeightApril26Blocks7_6Clean)
Survival_HeightApril26Blocks7_6Clean <-
Survival_HeightApril26Blocks7_6Clean |> 
  mutate(Seedling_Ht_cm=as.numeric(Seedling_Ht_cm)) |> 
  mutate(Num_leaves=as.numeric(Num_leaves)) |> 
  mutate(Num_dead_leaves=as.numeric(Num_dead_leaves)) |> 
  mutate(Leaf_greenness=as.numeric(Leaf_greenness)) |> 
  mutate(Num_germinated=as.numeric(Num_germinated)) |> 
  mutate(RCD=as.numeric(RCD))



```

```{r}


```

### Import remaining Oak files

```{r}
#importing the oak data files into session

oak_data <- list.files(path="Oak Project/Data",
                       
                       pattern="*6Clean.xlsx",
                       
                       full.names=TRUE) |> 
  
  lapply(read_excel,
         
         na="NA",sheet="Input1", skip=1 ) |> 
  bind_rows()

oak_data

view(oak_data)
```

### Rename Columns of Oak files

```{r}
oak_data <-
  oak_data |> 
  dplyr::rename(Page_Num ="Page #",
                Date ="MM/DD/YYYY", 
                Recorder ="Name",
                Block  ="#...4",
                Tag_Num   ="#...5",
                Oak_type  ="NR, S, or A",
                Treatment ="AC, OA, NP, DT, or DL",
                OA_microclimate    ="TH, TI, CN",
                Seedling_Ht_cm  ="(cm)",
                Num_leaves ="#...10",
                Num_dead_leaves  ="#...11",
                Leaf_greenness ="%",
                Num_germinated ="#...13",
                RCD  = "(mm)",
                Notes  ="as written on data sheet"
  )
```

```{r}
glimpse(oak_data)
oak_data <-
  oak_data |> 
  mutate(Seedling_Ht_cm=as.numeric(Seedling_Ht_cm)) |> 
  mutate(Num_leaves=as.numeric(Num_leaves)) |> 
  mutate(Num_dead_leaves=as.numeric(Num_dead_leaves)) |> 
  mutate(Leaf_greenness=as.numeric(Leaf_greenness)) |> 
  mutate(Num_germinated=as.numeric(Num_germinated)) |> 
  mutate(RCD=as.numeric(RCD))


```

```{r}
#splitting tag number and acorn number into separate columns
oak_data_CLEAN<- 
  oak_data |> 
  separate(Tag_Num, into = c("Tag_Num", "Acorn_Num"), sep="A") |> 
  mutate(Tag_Num=as.numeric(Tag_Num),
         Acorn_Num=as.numeric(Acorn_Num)) |> 
  mutate(Leaf_greenness=as.numeric(Leaf_greenness),
         Dead = case_when(
           Seedling_Ht_cm==0~"Dead",
           Seedling_Ht_cm>0~"Alive"
         ))

oak_data_CLEAN


```

```{r}
unloadNamespace("plyr")
#seedling height
seedling_summary <- 
  oak_data_CLEAN |>
  select(Treatment, OA_microclimate, Seedling_Ht_cm) |> 
  group_by(OA_microclimate) |> 
  summarize(
            n= n(),
            MeanSeedHeight = mean(Seedling_Ht_cm))

#ggplot(OA_microclimate)
#aes(x = Microclimate)
#geom_bar()
```

```{r}
oak_data_CLEAN |> 
  select(Tag_Num, Acorn_Num) |> 
  distinct() |> 
  arrange(Tag_Num)

#Acorn_ID_Table<- oak_data_CLEAN |> 
#    select(Tag_Num) |> 
#    arrange(Tag_Num) |> 
#  mutate(Acorn1=1, Acorn2=2, Acorn3=3) |> 
#  pivot_longer(
#    cols=c(Acorn1, Acorn2, Acorn3),
#    names_to = "Whatever",
#    values_to = "Acorn_Num" |> 
#      select(-Whatever))
```

Q. Effect of treatment:

All 4 responses:

-Seedling height

- focus on focal

- acorn vs NR analyze separate

  - acorn: at end

  - NR: over time

- exclude NA

-Seedling greenness

- focus on focal

- acorn vs NR analyze separate

  - acorn: at end

  - NR: over time

- exclude NA

-germination rate: each tag once, last time only, exclude NRs, only Acorns

-seedling survival: over time OR at end, exclude NAs (non germinates), ?clean data more, summarize to RATE OR logistic regression

!!!!Next assignment: Germination rate for microclimate

### Germination Rate

```{r}

oak_data_CLEAN |> 
  select(OA_microclimate)

```
